Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
edit curves with brian's new edits, veronese in P^n

\ssec{Effective Divisors of Hypersurfaces on Projective Space}
\label{ssec:proj-one-point}

\begin{rem}
\label{rem:proj-veronese}
For any divisor $D = \sum_{i = 0}^{n} \alpha_i C_i$ on $\bp^m$
with each $C_i$ a hypersurface, we can use the Veronese mappings
to view $D$ as a divisor on $\bp^{m'}$ for some $m' \in \bz_{\geq 0}$
with $D = \sum_{i = 0}^{n} \alpha_i H_i$ where $H_i$ are hypersurfaces
in $\bp^{m'}$.
\end{rem}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
Note that the method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}

@@

\begin{proof}
First, we apply Remark ~\ref{rem:proj-veronese} to reduce the
problem to letting each $H_i$ be a hyperplane of $\bp^m$.
Then, the proof follows from applying Remark
~\ref{rem:proj-one-point-ind} inductively on a one-hyperplane case
from Theorem ~\ref{thm:proj-one-point}.
\end{proof}

>>>>

\ssec{Effective Divisors of Hypersurfaces on Projective Space}
\label{ssec:proj-one-point}
\todo{intro}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
The method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}

\begin{rem}
\label{rem:proj-veronese}
For any divisor $D = \sum_{i = 0}^{n} \alpha_i C_i$ on $\bp^m$
with each $C_i$ a hypersurface and $\alpha_i \in \br$, we can use
the Veronese mappings to view $D$ as a divisor on $\bp^{m'}$ for some $m' \in \bz_{\geq 0}$
with $D = \sum_{i = 0}^{n} \alpha_i H_i$ where $H_i$ are 
hypersurfaces
in $\bp^{m'}$.
\end{rem}

@@

\begin{proof}
We derive the result from the following inductive argument.
If $n = 0$, i.e.~ $D = 0$, then we are done.
Now, we look to inductively add hypersurfaces.
If $D' = \sum_{i = 0}^{n} \alpha_i C_i$ is some $\bq$-divisor
on $\bp^m$, then we look at $D = D' + C$ where $C$ is a
hypersurface of degree $\pdeg$. Now use the Veronese mapping
to view $D$ as a divisor on $\bp^{m ({m + \pdeg} \choose \pdeg - 1)}$
and the image of $C$ as the intersection of a Veronese variety
with a hyperplane. Then we can reduce inductively adding hypersurfaces
to the case of adding hyperplanes. However, this is accomplished
by the method used in Theorem ~\ref{thm:proj-one-point} in an
inductive manner as mentioned in Remark ~\ref{rem:proj-one-point-ind}.
\end{proof}
